NDNS multilevel LCA CW2CB2 or CW2CB3

### 2018-08-10 paper/Chapters/Chapter2.tex
WCtab2 done describe

\subsection{Association between individual level latent classes and the prevalence of hypertension, obesity, and diabetes}\vspace{-0.3cm}

Associations between individual level carbohydrate eating classes and the prevalence of hypertension, obesity as well as diabetes were explored in men and women separately. Point estimates of weighted means and proportions and 95\%CI of the characteristics were determined by applying either nurse visiting weights (for outcomes of hypertension, BMI, and WC) or blood sample weights (for outcomes of DM) accordingly. Similarly, the \textit{F} test (for continuous variables) and adjusted Pearson $\chi^2$ test (for categorical variables) were used to determine sex-specific differences by hypertension status, BMI categories, and DM diagnosis.

@@

Logistic regression models (for hypertension, and diabetes), and linear regression model (for WC, BMI), were used to test for associations between individual level carbohydrate eating classes and hypertension, diabetes, obesity, and WC, separately.

>>>>

\subsection{Association between individual level latent classes and the prevalence of hypertension, and measurements of obesity}\vspace{-0.3cm}

Associations between individual level carbohydrate eating classes and hypertension (yes/no), body mass index (BMI, kg/m\textsuperscript{2}), and waist circumference (WC, cm) were explored in men and women separately. Point estimates of weighted means and proportions and 95\%CI of the characteristics were determined by applying either nurse visiting weights (for outcomes of hypertension, BMI, and WC) or blood sample weights (for diagnosis of DM) accordingly. Similarly, \textit{F} tests (for continuous variables) and adjusted Pearson $\chi^2$ tests (for categorical variables) were used to determine sex-specific differences by hypertension status, and BMI categories.

@@

Survey-designed logistic regression models (for hypertension), and linear regression models (for WC, BMI), were used to test for associations between latent classes of carbohydrate eating patterns and hypertension, BMI, and WC, in the NDNS RP sample, separately. Since diabetic participants might or might not modify their carbohydrate eating habits, we also conducted all the above mentioned regression models restricted to those without diabetes.

### 2018-08-10 paper/Chapters/Chapter3.tex
WCtab2 done describe

Results of the regression analyses showed inverse associations between latent classes of carbohydrate eating patterns and BMI among men \textbf{(Table \ref{tab:tab2BMI})}. However, the 95\%CI of the regression coefficients were all too wide and include the null value 0, indicating no statistically supported evidence of the association. But, evidence of interaction effect was found in whether live a partner or not on the association between carbohydrate eating patterns and BMI (\textit{p} for interaction = 0.014 and 0.036 for women in total and women without diabetes). For women living with her partner, latent classes of carbohydrate eating patterns were negatively associated with BMI. Compared with women eating low carbohydrate meals, women having

@@

	\caption{\label{tab:tab2WC}Associations of carbohydrate eating patterns with BMI in the UK adults , with or without diabetes. \\ (NDNS RP 2008/09-15/16, sample size = 6155)}\vspace{-0.3cm}
	\centering
	\fontsize{9}{11}\selectfont
	\begin{tabular}[t]{lccccc}
		\hiderowcolors
		\toprule
		\multicolumn{1}{c}{ } & \multicolumn{5}{c}{\textbf{Carbohydrate eating patterns}} \\
		\cmidrule(l{2pt}r{2pt}){2-6}
		\textbf{Model} & \textbf{Low} & \textbf{Moderate} & \textbf{\textit{P} value\textsuperscript{*}} & \textbf{High} & \textbf{\textit{P} value\textsuperscript{*}}\\
		\midrule
		\showrowcolors
		\addlinespace[0.3em]
		\multicolumn{6}{l}{\textbf{Men (n = 2537)}}\\
		\addlinespace[0.3em]
		\multicolumn{6}{l}{\hspace{1em}\textbf{WC}}\\
		\hspace{1em}\hspace{1em}Crude model & -- & -3.00 (-5.49, -0.52) & 0.018 & -0.90 (-2.84, 1.04) & 0.364\\
		\hspace{1em}\hspace{1em}Model 1\textsuperscript{\dag} & -- & 1.06 (-1.50, 3.64) & 0.415 & -1.55 (-3.42, 0.31) & 0.103\\
		\addlinespace[0.3em]
		\multicolumn{6}{l}{\hspace{1em}\textbf{WC in non-diabetics}}\\
		\hspace{1em}\hspace{1em}Crude model & -- & -2.51 (-5.00, -0.21) & 0.048 & -0.51 (-2.47, 1.44) & 0.606\\
		\hspace{1em}\hspace{1em}Model 1 & -- & 1.42 (-1.17, 4.01) & 0.283 & -1.29 (-3.18, 0.60) & 0.181\\
		\addlinespace[0.3em]
		\multicolumn{6}{l}{\textbf{Women (n = 3618)}}\\
		\addlinespace[0.3em]
		\multicolumn{6}{l}{\hspace{1em}\textbf{WC}}\\
		\hspace{1em}\hspace{1em}Crude model & -- & -1.28 (-3.26, 0.70) & 0.206 & -2.81 (-4.50, -1.12) & 0.001\\
		\hspace{1em}\hspace{1em}Live with partner\textsuperscript{\ddag} & -- & 0.28 (-2.85, 3.41) & 0.861 & -4.71 (-7.00, -2.43) & < 0.001\\
		\hspace{1em}\hspace{1em}Live alone & -- & 3.17 (0.05, 6.30) & 0.047 & 0.73 (-1.84, 3.30) & 0.577\\
		\addlinespace[0.3em]
		\multicolumn{6}{l}{\hspace{1em}\textbf{WC in non-diabetics}}\\
		\hspace{1em}\hspace{1em}Crude model & -- & -0.91 (-2.88, 1.07) & 0.368 & -2.41 (-4.06, -0.76) & 0.004\\
		\hspace{1em}\hspace{1em}Live with partner & -- & 1.11 (-2.02, 4.23) & 0.487 & -3.74 (-5.97, -1.51) & 0.001\\
		\hspace{1em}\hspace{1em}Live alone & -- & 3.08 (-0.09, 6.25) & 0.057 & 0.16 (-2.36, 2.69) & 0.899\\
		\bottomrule
		\multicolumn{6}{l}{{\scriptsize \textit{Note: }}}\\
		\multicolumn{6}{l}{{\scriptsize \textbf{Abbreviations}: WC, waist circumference; NDNS RP, national dietary and nutrition survey rolling programme.}}\\
		\multicolumn{6}{l}{{\scriptsize Diabetes was defined by A1C > 6.5\%. }}\\
		\multicolumn{6}{l}{{\scriptsize \textsuperscript{*} P values were obtained from wald tests from linear regression models.}}\\
		\multicolumn{6}{l}{{\scriptsize \textsuperscript{\dag} Adjusted for age, live with partner or not, average household income, education level, hypertension (yes or no),}}\\
		\multicolumn{6}{l}{{\scriptsize smoking status, total energy intake, alcohol consumption;}}\\
		\multicolumn{6}{l}{{\scriptsize \textsuperscript{\ddag} Adjusted for age, education level, hypertension, smoking status, total energy intake, alcohol consumption.}}\\
	\end{tabular}
\end{table}
\vspace{-0.3cm}
\rowcolors{2}{white}{white}

Similarly, when using WC as another measurement of obesity, compared with men who had low carbohydrate eating pattern, only those had a moderate carbohydrate eating pattern were found with about 3 cm smaller WC in the crude model \textbf{(Table \ref{tab:tab2WC})}. After adjustment for

>>>>

Results of the regression analyses showed inverse associations between latent classes of carbohydrate eating patterns and BMI among men \textbf{(Table \ref{tab:tab2BMI})}. However, the 95\%CI of the regression coefficients were all too wide and include the null value 0, indicating no statistically supported evidence of the association. But, evidence of interaction effect was found in whether live with a partner or not on the association between carbohydrate eating patterns and BMI (\textit{p} for interaction = 0.014 and 0.036 for women in total and without diabetes). For women living with her partner, latent classes of carbohydrate eating patterns were negatively associated with BMI. Compared with women eating low carbohydrate meals, women having

@@

	\caption{\label{tab:tab2WC}Associations of carbohydrate eating patterns with BMI in the UK adults , with or without diabetes. \\ (NDNS RP 2008/09-15/16, sample size = 6155)}\vspace{-0.3cm}
	\centering
	\fontsize{9}{11}\selectfont
	\begin{tabular}[t]{lccccc}
		\hiderowcolors
		\toprule
		\multicolumn{1}{c}{ } & \multicolumn{5}{c}{\textbf{Carbohydrate eating patterns}} \\
		\cmidrule(l{2pt}r{2pt}){2-6}
		\textbf{Model} & \textbf{Low} & \textbf{Moderate} & \textbf{\textit{P} value\textsuperscript{*}} & \textbf{High} & \textbf{\textit{P} value\textsuperscript{*}}\\
		\midrule
		\showrowcolors
		\addlinespace[0.3em]
		\multicolumn{6}{l}{\textbf{Men (n = 2537)}}\\
		\addlinespace[0.3em]
		\multicolumn{6}{l}{\hspace{1em}\textbf{WC}}\\
		\hspace{1em}\hspace{1em}Crude model & -- & -3.00 (-5.49, -0.52) & 0.018 & -0.90 (-2.84, 1.04) & 0.364\\
		\hspace{1em}\hspace{1em}Model 1\textsuperscript{\dag} & -- & 1.06 (-1.50, 3.64) & 0.415 & -1.55 (-3.42, 0.31) & 0.103\\
		\addlinespace[0.3em]
		\multicolumn{6}{l}{\hspace{1em}\textbf{WC in non-diabetics}}\\
		\hspace{1em}\hspace{1em}Crude model & -- & -2.51 (-5.00, -0.21) & 0.048 & -0.51 (-2.47, 1.44) & 0.606\\
		\hspace{1em}\hspace{1em}Model 1\textsuperscript{\dag} & -- & 1.42 (-1.17, 4.01) & 0.283 & -1.29 (-3.18, 0.60) & 0.181\\
		\addlinespace[0.3em]
		\multicolumn{6}{l}{\textbf{Women (n = 3618)}}\\
		\addlinespace[0.3em]
		\multicolumn{6}{l}{\hspace{1em}\textbf{WC}}\\
		\hspace{1em}\hspace{1em}Crude model & -- & -1.28 (-3.26, 0.70) & 0.206 & -2.81 (-4.50, -1.12) & 0.001\\
		\hspace{1em}\hspace{1em}Live with partner\textsuperscript{\ddag} & -- & 0.28 (-2.85, 3.41) & 0.861 & -4.71 (-7.00, -2.43) & < 0.001\\
		\hspace{1em}\hspace{1em}Live alone\textsuperscript{\ddag} & -- & 3.17 (0.05, 6.30) & 0.047 & 0.73 (-1.84, 3.30) & 0.577\\
		\addlinespace[0.3em]
		\multicolumn{6}{l}{\hspace{1em}\textbf{WC in non-diabetics}}\\
		\hspace{1em}\hspace{1em}Crude model & -- & -0.91 (-2.88, 1.07) & 0.368 & -2.41 (-4.06, -0.76) & 0.004\\
		\hspace{1em}\hspace{1em}Live with partner\textsuperscript{\ddag} & -- & 1.11 (-2.02, 4.23) & 0.487 & -3.74 (-5.97, -1.51) & 0.001\\
		\hspace{1em}\hspace{1em}Live alone\textsuperscript{\ddag} & -- & 3.08 (-0.09, 6.25) & 0.057 & 0.16 (-2.36, 2.69) & 0.899\\
		\bottomrule
		\multicolumn{6}{l}{{\scriptsize \textit{Note: }}}\\
		\multicolumn{6}{l}{{\scriptsize \textbf{Abbreviations}: WC, waist circumference; NDNS RP, national dietary and nutrition survey rolling programme.}}\\
		\multicolumn{6}{l}{{\scriptsize Diabetes was defined by A1C > 6.5\%. }}\\
		\multicolumn{6}{l}{{\scriptsize \textsuperscript{*} P values were obtained from wald tests from linear regression models.}}\\
		\multicolumn{6}{l}{{\scriptsize \textsuperscript{\dag} Adjusted for age, live with partner or not, average household income, education level, hypertension (yes or no),}}\\
		\multicolumn{6}{l}{{\scriptsize smoking status, total energy intake, alcohol consumption;}}\\
		\multicolumn{6}{l}{{\scriptsize \textsuperscript{\ddag} Adjusted for age, education level, hypertension, smoking status, total energy intake, alcohol consumption.}}\\
	\end{tabular}
\end{table}
\vspace{-0.3cm}
\rowcolors{2}{white}{white}

Similarly, when using WC as another measurement of obesity, compared with men who had low carbohydrate eating pattern, only those had a moderate carbohydrate eating pattern were found with about 3 cm (95\%CI: 0.52, 5.49 cm, \textit{p} = 0.018) smaller WC in the crude model \textbf{(Table \ref{tab:tab2WC})}. After adjustment of age, living with partner or not, average household income, education level, hypertension, smoking status, total energy intake, and alcohol consumption, the association attenuated to no difference in men. Again, the interaction effect of whether living with partner or not on the association between carbohydrate eating patterns and WC was found in women in total and without diabetes. Among women who were living with her partner, high carbohydrate eaters had 4.71 cm (95\%CI: 2.43, 7.00, \textit{p} < 0.001) lower WC on average compared with low carbohydrate eaters. The association remained when restricted into non-diabetic women (-3.74 cm, 95\% CI: 1.51, 5.97, \textit{p} = 0.001). However, for women who lived without a partner, moderate carbohydrate eaters had 3.17 cm (95\%CI: 0.05, 6.30, \textit{p} = 0.043) larger WC compared with low carbohydrate eaters. The evidence for the positive association between moderate carbohydrate eaters and WC was weak and became borderline significant when excluding diabetic women.
